---
title: "Vector Autoregression Dashboard"
author: "David A. Bignotti"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    theme: lumen
    logo: 'C:/Users/File/Path/to/logo.png'
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(DT)
library(forecast)
library(vars)
library(tseries)
library(xts)
library(readxl)
library(fredr)
library(RColorBrewer)
library(MASS)
library(zoo)
library(urca)
library(astsa)
library(lmtest)
library(data.table)
library(htmltools)
library(dygraphs)
library(timeSeries)
library(cowplot)
```

```{r monthly update objects, include=FALSE}
# Update these objects each month to reflect the lastest input window and output window

input_start_date <- "2015-01-01" # the start date of input data
input_end_date <- "2019-07-01" # the end date of input data
output_start_date <- "2019-08-01" # the first month of 6-month forecast
output_end_date <- "2020-01-01" # the last month of 6-month forecast
dy_range_start_date <- "2018-01-01" # the first month to display in the forecast plot's range selector tool
```

```{r import and subset data, include=FALSE}
# Import the dataset from its local filepath

var_inputs <- read_excel("C:/Users/File/Path/to/input/data.xlsx")

fredr_set_key("APINUMBER") # API call to fred.stlouisfed.org

gasprice <- fredr(series_id = "GASREGM", 
                  observation_start = as.Date(input_start_date), 
                  observation_end = as.Date(input_end_date)
                  ) # bring in gas prices as an input variable

# Subset the input input data

gmc <- dplyr::filter(var_inputs, Division == "GMC")
gmc <- bind_cols(gmc, gasprice) %>% # combine the GMC and gas price data columns
  dplyr::select(-Month, -Year, -Division, -date, -series_id) # remove those columns not needed for model
gmc <- rename(gmc, GasPrice = value) # rename the gas price column
setcolorder(gmc, c("StartDate", "GasPrice", "TotalOutcomes", "ActualSales")) # reorder the columns so Sales is on the far right
gmc$StartDate <- as.Date(gmc$StartDate)
```

```{r VAR model, include=FALSE}
var <- dplyr::select(buick, ActualSales, TotalOutcomes, GasPrice) %>%
  VAR(lag.max = 6, type = "const", ic = c("AIC", "HQ", "SC", "FPE"))
```

```{r Impulse Response Function, include=FALSE}
irf1 <- vars::irf(x = var, 
                  impulse = "TotalOutcomes", 
                  response = "ActualSales", 
                  n.ahead = 6, 
                  ortho = TRUE, 
                  cumulative = FALSE, 
                  boot = TRUE, 
                  ci = 0.95)
irf2 <- vars::irf(x = var, 
                  impulse = "TotalOutcomes", 
                  response = "ActualSales", 
                  n.ahead = 6, 
                  ortho = TRUE, 
                  cumulative = TRUE, 
                  boot = TRUE, 
                  ci = 0.95)
```

```{r BUICK VAR forecast, include=FALSE}
# create a date field to join with forecasts
StartDate <- seq(from = as.Date(output_start_date), to = as.Date(output_end_date), by = 'month') 


# produce a six-month-ahead forecast with a 95% confidence interval and convert to data.frame
forecast <- predict(var, n.ahead = 6, ci = 0.95) 
fcast <- as.data.frame(forecast$fcst)


# reorder variables and keep only Leads and Sales forecasts
fcast <- fcast[, c(6, 5, 7, 2, 1, 3)] 


# combine forecast with dates
fcast <- cbind(StartDate, fcast) 


# split forecast data.frame into one each for leads and sales
fcastsales <- fcast[c(1, 5:7)]
fcastleads <- fcast[c(1, 2:4)]


# rename fcast values to agree with buick.past names
colnames(fcastsales)[3] <- "ForecastSales"
colnames(fcastleads)[3] <- "ForecastLeads"


# create prior values data.frames to join with forecast data.frames
pastsales <- gmc %>%
  dplyr::select(-GasPrice, -TotalOutcomes)

# create placeholder columns for confidence interval bounds

pastsales$ActualSales.lower <- NA  
pastsales$ActualSales.upper <- NA


# reorder columns 
pastsales <- pastsales[, c(1, 3, 2, 4)] 


# set placeholders as numeric field for joining
pastsales$ActualSales.lower <- as.numeric(pastsales$ActualSales.lower) 
pastsales$ActualSales.upper <- as.numeric(pastsales$ActualSales.upper)


# Join past data.frame with forecast data.frame
allsales <- merge(pastsales, fcastsales, all = TRUE)


# create prior values data.frames to join with forecast data.frames
pastleads <- gmc %>%
  dplyr::select(-GasPrice, -ActualSales)


# create placeholder columns for confidence interval bounds
pastleads$TotalOutcomes.lower <- NA
pastleads$TotalOutcomes.upper <- NA


# reorder columns 
pastleads <- pastleads[, c(1, 3, 2, 4)]


# set placeholders as numeric field for joining
pastleads$TotalOutcomes.lower <- as.numeric(pastleads$TotalOutcomes.lower) 
pastleads$TotalOutcomes.upper <- as.numeric(pastleads$TotalOutcomes.upper)


# Join past data.frame with forecast data.frame
allleads <- merge(pastleads, fcastleads, all = TRUE)


# remove date field and transform sales data to time series objects (in order to function with dygraphs plot)
sales <- allsales %>%
  dplyr::select(-StartDate) 


# transform sales data.frames into time series objects
sales <- ts(sales, start = c(2015, 1), frequency = 12)


# remove date field and transform leads data to time series objects (in order to function with dygraphs plot)
leads <- allleads %>%
  dplyr::select(-StartDate)


# transform leads data.frames into time series objects
leads <- ts(leads, start = c(2015, 1), frequency = 12)


# remove objects to clear local memory
rm(StartDate)
rm(fcastleads)
rm(fcastsales)
rm(pastleads)
rm(pastsales)
rm(allleads)
rm(allsales)
rm(sales)
rm(leads)
```


Sidebar {.sidebar data-width=250}
=====================================



**Navigating this Dashboard**

* **Historic Data** tracks the average national price per gallon of gasoline, reported lead outcomes for the brand, and retail deliveries for every month dating back to 2015.

* **Impulse Response Functions** approximate the influence of a 1% spike in lead outcomes on retail deliveries, as far as six months after the fact. In this panel, month 1 refers to the month in which a theoretical 1% (positive) jump in lead outcomes occurs. Month 2 refers to the next month's expected change in retail deliveries attributable to the spike in lead outcomes.

* **Forecast Data** shows the six-month forecast for lead outcomes and retail sales, including a 95% confidence interval. We expect, with 95% certainty, the number of lead outcomes and retail sales to fall within the lower and upper bounds of our forecast.

* **Forecast Plots** show the historic leads and sales data and the forecasted values for the next six months.

* For questions/comments, please contact David Bignotti (email@email.com)


GMC {data-icon="fa-expand"}
=====================================  

Column {data-width=375}
-----------------------------------------------------------------------

### GMC Historic Data {data-height=285}

```{r Buick Historic Values Table Panel}
datatable(gmc,
          rownames = FALSE,
          colnames = c('Month', 'Gas Price', 'Lead Outcomes', 'Retail Deliveries'),
          extensions = c('Scroller', 'Buttons', 'Responsive'),
          class = 'compact table-hover',
          options = list(dom = 'rt',
                         buttons = c('copy', 'csv', 'excel'),
                         pageLength = 12,
                         scrollY = "200",
                         scroller = TRUE,
                         columnDefs = list(list(className = 'dt-center',
                                                targets = 1:3), 
                                           list(className = 'dt-left', 
                                                targets = 0)))) %>%
  formatRound('TotalOutcomes', interval = 3, mark = ",", digits = 0) %>%
  formatRound('ActualSales', interval = 3, mark = ",", digits = 0) %>%
  formatCurrency('GasPrice', currency = "$", digits = 2)
```

### {data-height=365}

```{r Impulse Response Functions, echo=FALSE, out.width= "50%"}
f1 <- plot(irf1$irf$TotalOutcomes, 
             type = "n", 
             ylab = "Retail Deliveries", 
             sub = "95% Bootstrap Confidence Interval", 
             main = "Impulse Response from Lead Outcomes", 
             ylim = c(min(irf1$Lower$TotalOutcomes), max(irf1$Upper$TotalOutcomes))) 
abline(h = 0, 
       lwd = 2) 
polygon(x = c(seq_along(irf1$irf$TotalOutcomes), rev(seq_along(irf1$irf$TotalOutcomes))), 
        y = c(irf1$Lower$TotalOutcomes, rev(irf1$Upper$TotalOutcomes)), 
        lty = 0, 
        col = adjustcolor("lightgrey", alpha.f = 0.5)) 
lines(irf1$irf$TotalOutcomes, 
      col = "steelblue2", 
      lwd = 5)

f2 <- plot(irf2$irf$TotalOutcomes, 
             type = "n", 
             ylab = "Retail Deliveries", 
             sub = "95% Bootstrap Confidence Interval", 
             main = "Cumulative Impulse Response from Lead Outcomes", 
             ylim = c(min(irf2$Lower$TotalOutcomes), max(irf2$Upper$TotalOutcomes)))
abline(h = 0, 
       lwd = 2)
polygon(x = c(seq_along(irf2$irf$TotalOutcomes), rev(seq_along(irf2$irf$TotalOutcomes))), 
        y = c(irf2$Lower$TotalOutcomes, rev(irf2$Upper$TotalOutcomes)), 
        lty = 0, 
        col = adjustcolor("lightgrey", alpha.f = 0.5))
lines(irf2$irf$TotalOutcomes, 
      col = "steelblue2", 
      lwd = 5)

cowplot::plot_grid(f1, f2, align = "h", nrow = 1, ncol = 2, labels = NULL)
```

### GMC 6-Month Sales Forecast {data-height=350}

```{r Buick Forecast Values Table Panel}
fcast <- fcast %>%
  dplyr::select(-StartDate)

custom <- htmltools::withTags(table(
  class = 'display', 
  thead(
    tr(
      th(rowspan = 2, 'Months Ahead'),
      th(colspan = 3, 'Lead Outcomes'),
      th(colspan = 3, 'Retail Deliveries')
    ),
    tr(
      lapply(rep(c('Lower Bound', 'Forecast', 'Upper Bound'), 2), th)
    )
  )
))

datatable(fcast, 
          rownames = TRUE, 
          class = 'compact table-hover', 
          container = custom, 
          colnames = c('Leads Lower Bound', 'Leads Forecast', 'Leads Upper Bound', 
                       'Sales Lower Bound', 'Sales Forecast', 'Sales Upper Bound'), 
          extensions = c('Buttons', 'Responsive'), 
          options = list(dom = 'rt', 
                         buttons = c('copy', 'csv', 'excel'), 
                         pageLength = 6, 
                         columnDefs = list(list(className = 'dt-center', 
                                                targets = 0:6))))  %>%
  formatRound(1:6, interval = 3, digits = 0)
```


Column {data-width=550}
-----------------------------------------------------------------------

###

```{r GMC Sales Forecast Plot}
dygraph(sales, main = "GMC Retail Sales Forecast") %>%
  dySeries("ActualSales", 
           label = "Actual", 
           strokeWidth = 2, 
           color = "green") %>%
  dySeries(c("ActualSales.lower", "ForecastSales", "ActualSales.upper"), 
           label = "Forecast", 
           strokeWidth = 2, 
           strokePattern = "dashed", 
           drawPoints = TRUE, 
           color = "#7f0180") %>%
  dyAxis("y", label = "Units Delivered", 
         valueRange = c(0, upperbound), 
         axisLineWidth = 1.5, 
         drawGrid = TRUE, 
         axisLabelWidth = 70) %>%
  dyAxis("x", label = "Month", 
         drawGrid = TRUE, 
         axisLineWidth = 1.5) %>%
  dyEvent(output_start_date, 
          "Forecast Start     ", 
          labelLoc = "top") %>% 
  dyHighlight(highlightCircleSize = 5,
              highlightSeriesBackgroundAlpha = 0.5,
              highlightSeriesOpts = list(strokeWidth = 3),
              hideOnMouseOut = FALSE) %>%
  dyOptions(rightGap = 20, 
            labelsKMB = TRUE) %>%
  dyRangeSelector(height = 20, 
                  dateWindow = c(dy_range_start_date, output_end_date))
```

###

```{r GMC Leads Forecast Plot}
dygraph(leads, main = "Buick Lead Outcomes Forecast") %>%
  dySeries("TotalOutcomes", 
           label = "Actual", 
           strokeWidth = 2, 
           color = "green") %>%
  dySeries(c("TotalOutcomes.lower", "ForecastLeads", "TotalOutcomes.upper"), 
           label = "Forecast", 
           strokeWidth = 2, 
           strokePattern = "dashed", 
           drawPoints = TRUE, 
           color = "#7f0180") %>%
  dyAxis("y", label = "Outcomes Generated", 
         valueRange = c(0, upperbound), 
         axisLineWidth = 1.5, 
         drawGrid = TRUE, 
         axisLabelWidth = 70) %>%
  dyAxis("x", label = "Month", 
         drawGrid = TRUE, 
         axisLineWidth = 1.5) %>%
  dyEvent(output_start_date, # MANUAL UPDATE
          "Forecast Start     ", 
          labelLoc = "top") %>% 
  dyHighlight(highlightCircleSize = 5,
              highlightSeriesBackgroundAlpha = 0.5,
              highlightSeriesOpts = list(strokeWidth = 3),
              hideOnMouseOut = FALSE) %>%
  dyOptions(rightGap = 20, 
            labelsKMB = TRUE) %>%
  dyRangeSelector(height = 20, 
                  dateWindow = c(dy_range_start_date, output_end_date))
```
